\textbf{ Mapping stellar surfaces\\ I: Degeneracies in the light curve problem }

### 2021-01-22 paper1/ms.tex
more text

It is instructive to actually visualize these components in a
an actual surface mapping exercise. Figure~\ref{fig:nullspace}
shows the decomposition of four hypothetical surfaces (left
column) into its preimage (center) and the null space (right)
components under the flux operator $\mathbf{A}$, which we compute for
definiteness at an inclination of $I = 60^\circ$ over a full rotation.
The light curves corresponding to each of these surfaces are also shown.
Note that both the true map and its associated light curve are simply equal to
the sum of the preimage and null space components.%
\footnote{%
    This follows from the orthogonality of $\mathbf{V}$:
    $
        \mathbf{V} \mathbf{V}^\top =
        \mathbf{I} =
        \mathbf{V}_\bullet \mathbf{V}_\bullet^\top  + \mathbf{V}_\circ \mathbf{V}_\circ^\top =
        \mathbf{P} + \mathbf{N}
    $, so $\mathbf{y} = \mathbf{y}_\bullet + \mathbf{y}_\circ$.
}

As expected, all of the information in the light curve comes
from the preimage, and the null space contributes exactly zero
flux at all phases. However, most of the information about the
\emph{surface} is stuck in the null space!

\xxx{More on this.}

@@

\subsection{Why ensemble analyses work}
\label{sec:ensemble}

>>>>

It is instructive to actually visualize these components in a
an actual surface mapping exercise. Figure~\ref{fig:nullspace}
shows the decomposition of four hypothetical surfaces (left
column) into its preimage (center) and the null space (right)
components under the flux operator $\mathbf{A}$, which we compute for
definiteness at an inclination of $I = 60^\circ$ over a full rotation.
The light curves corresponding to each of these surfaces are also shown.
Note that both the true map and its associated light curve are simply equal to
the sum of the preimage and null space components.%
\footnote{%
    This follows from the orthogonality of $\mathbf{V}$:
    $
        \mathbf{V} \mathbf{V}^\top =
        \mathbf{I} =
        \mathbf{V}_\bullet \mathbf{V}_\bullet^\top  + \mathbf{V}_\circ \mathbf{V}_\circ^\top =
        \mathbf{P} + \mathbf{N}
    $, so $\mathbf{y} = \mathbf{y}_\bullet + \mathbf{y}_\circ$.
}
%
As expected, all of the information in the light curve comes
from the preimage, and the null space contributes exactly zero
flux at all phases. However, most of the information about the
\emph{surface} is stuck in the null space!

In the top panel of the figure, it is clear that the light curve
contains \emph{some} information about the presence of a spot at roughly the
correct longitude and latitude. However, there are spurious features

\xxx{More on this.... Discuss each of the rows in the Figure...}

It may be helpful to consider Figure~\ref{fig:nullspace} in the
context of an inference exercise. One can equivalently think of
the preimage as the solution to a least squares inference problem
in which a completely uninformative prior%
\footnote{%
    Specifically, we place a zero-mean, infinite variance prior on the coefficients of $\mathbf{y}$.
}
is placed on the spherical harmonic
coefficients. In this case, the least squares solution is given by
%
\begin{proof}{test_svd}
    \hat{\mathbf{y}}
    &=
    \lim_{\lambda \rightarrow 0} \left( \mathbf{A}^\top \mathbf{A} + \lambda \mathbf{I}\right)^{-1} \mathbf{A}^\top \mathbf{f}
    \nonumber\\
    &= \mathbf{y}_\bullet
    \quad.
\end{proof}
%
In this sense, $\mathbf{y}_\bullet$ represents our knowledge about the
surface of a star after an observation if we have no prior information
whatsoever on $\mathbf{y}$.

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=0.5\linewidth]{figures/infinite_pixels.pdf}
        \oscaption{infinite_pixels}{%
            Foo.
            \label{fig:infinite_pixels}
        }
    \end{centering}
\end{figure}

Another point to keep in mind is that the particular decomposition of the
surface map into its preimage and its nullspace will depend on the basis
if the basis is not \emph{complete} (which, in practice, will almost
certainly be the case).%

\xxx{FOO BAR}

%
In other words, if we instead carry out our inference in a pixel basis
with a finite number of pixels $P$,
the preimage (or, equivalently, our least squares solution with an uninformative
prior) will be different than that shown in Figure~\ref{fig:nullspace}.
Interestingly, however,
this difference will tend to zero as $P$ is increased.
%
Figure~\ref{fig:infinite_pixels} shows the fractional difference between
the least squares solution to the mapping problem in the spherical harmonic
basis and the solution in the pixel basis, as a function of the number of
pixels $P$. The difference is computed in the spherical harmonic basis
by transforming the pixel solution
to the spherical harmonic basis (blue) and in the pixel basis by transforming
the spherical harmonic solution to the pixel basis (orange). Both methods yield
essentially the same result: the difference between the two solutions
tends to zero as the number of pixels $P$ is increased.%

\subsection{Dependence on inclination}

\xxx{Words here, Figure~\ref{fig:nullspace_ensemble}, etc.}

%
\begin{align}
    \label{eq:shrinkage}
    R \equiv 1 - \lim\limits_{\sigma_0^2 \rightarrow \infty}
    \frac{\sigma^2}{\sigma_0^2}
    \quad,
\end{align}
%

@@

\subsection{Ensemble analyses work}
\label{sec:ensemble}
